\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-12 intro.tex
update

>>>>

Autonomous platforms in the marine environment have had a substantial
and immediate impact for both civil and military
applications. Autonomous Underwater Vehicles (AUVs) in particular have
altered the way oceanographers are able to reach beyond the surface
and make observations which are not only at higher resolution but have
altered both the discipline and impacted the engineering methodologies
behind Sampling, Control and Robotics itself \comment{need higher
  level citation here}. Yet challenges remain, principally in making
such robotic devices more attune to the environment, to sample large
swaths at the Meso-scale ($> 50$ km\textsuperscript{2}) and doing so
systematically and adaptively. Further, over the last two decades
substantial improvements in hardware including propulsion, battery
technologies and hardware have not been kept abreast of in software
especially in algorithmic methods in Sampling and Control and in
systematic engineering and software engineering development.

This is one important reason that periodic calls towards sustained
exploration and sampling \cite{rudnick03} are issued by scientists for
whom the scope and scale of observations have become increasingly
tractable using autonomous platforms both below and on the
surface. The complexity of oceanic processes and the chronic
under-sampling of the coastal oceans especially of dynamic features
spread over large spatial scales (from meters to kilometers in extent)
with dynamic biological activity across the temporal spectrum (in the
order of hours to weeks), has made it challenging for robotic
platforms to autonomously track and sample. A prominent example of
such a process is coastal algal blooms which are patchy and could
cover large coastal zones. Persistent observation of such dynamic
events dictates that our robotic assets track and sample such patches
which can evolve rapidly due to inherent bio-geochemical activity,
advection as well as diffusion with the water mass it resides in.

At the Monterey Bay Aquarium Research Institute (MBARI), a unique
inter-disciplinary collaboration between biologists, ecologists,
geneticists and roboticists on the CANON (Controlled Agile Novel
Observation Network) \cite{canon} has driven autonomous system
exploration and control. Fundamental questions related to sampling
strategies have driven the use cases for advanced concepts in
autonomy, specifically to Artificial Intelligence (AI) Planning and
Execution. Current work is also looking at how other fields in Machine
Intelligence especially Machine Learning can be leveraged to impact
the science of engineering systems in marine robotics and as a
consequence oceanography itself.

In this chapter, we propose an alternative view to decision-making
using in-situ automated Planning \cite{ghallab04} and more important
demonstrate that while earlier control paradigms primarily driven by
\emph{reactive} approaches to control can be augmented to not only
provide targeted observation and sampling, but also a more nuanced and
balanced consideration of mission objectives, environmental conditions
and available resources. Planning (or deliberation or projection in
action space) is important to balance current needs of a robot with
future desires or goals. And to do such planning in the context of
time and available resources in-situ on the robot. Planning without
time or action planning in of itself is not adequate; dealing with
time ensures that the balance between now and the future is handled
systematically. Yet, plans have to be tied to robotic action and in
the real-world which is not static. Further, plans project future
state which is not only likely to change, but is based on substantial
uncertainty\footnote{It is important to remember that these principles
  are broad much as Dwight Eisenhower is reputed to have said
  ``\emph{In preparing for battle I have always found that plans are
    useless, but planning is indispensable}'' and ``\emph{Failing to
    plan is planning to fail}''}. Modeling of the marine environment
to provide cues is hampered by poor understanding of the worlds ocean
and poor predictive skill of ocean models and often little to no
availability of a priori synoptic views of the survey area.

AUV control architectures have primarily been driven by reactive
Subsumption-based approaches \cite{brooks86}. They have been adequate
to doing routine straight-line survey tracts which have played a
valuable role in the acceptance of AUVs as a new tool for ocean
science and DoD applications. However, scaling of problem domains in
addition to dealing with event-response and discovery in the
water-column challenges \emph{how} controllers are written in the
first place, maintained and upgraded over the life-cycle of the
platform. Dealing with off-nominal endogenous or exogenous conditions
including sensor failures stretches how AUVs are currently being
utilized; one consequence for instance, is any off-nominal condition
results in a \emph{fail-safe} surface mode to radio for help. Further
these approaches, words to the contrary are not only hard to maintain,
are not 'adaptive' enough. Despite vigorous defense by Brooks early on
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
require a model if not of the world then of the \emph{physics} of the
vehicle when it interacts with the world.

Persistent and early legitimate criticism of deliberation on the other
hand also well articulated by Brooks
\cite{Brooks91intelligencewithoutrea,Brooks91intelligencewithoutrep}
was targeting the performance of in-situ deliberation. Terrestrial
robotic platforms were well-known to frequently stop and ``plan''. In
the marine robotics world, such a situation would simply not be
permissible given the need to make continuous observations without
stopping an under-articulated robotic platform such as an AUV. However
algorithmic and implementational advances by the AI community over the
years has produced a range of planners with demonstrated embedded
capabilities \comment{cite}. Our work has been influenced by and in
turn influenced a wide range of activities in the field of AI Planning
and Plan Execution with demonstrated real-world and highly visible
capabilities in the Space domain
\cite{mus98,rajan00,ai-chang04,bresina05} for NASA. Lessons learned
have been applied to the world of marine robotics which we bring
together in this chapter. Important lessons that were derived from
this rich legacy involved the following:

\begin{itemize}

\item Projecting plans (or using \emph{deliberation}) can and should
  balance near-term objectives with end-term or even evolving goals.

\item Dealing with time and resources is essential in real-world
  planning.

\item A rich representation that can deal with co-temporal events for
  modeling complex systems is a must.

\item Fast solvers for incremental causal reasoning can be built as a
  basis for dealing with dynamic controllable events where replanning
  is tightly integrated.

\item Planning should be tightly inter-leaved with execution to enable
  responsiveness to events which impact plan failure. This is also an
  important factor leading to platform adaptivity.

\item A model-based approach separating control formulation of the
  platform which describes operational characteristics from the
  domain. The controller can then be rigorously tested and certified
  even as the domain model evolves to fit diverse applications.

\end{itemize}

We have developed, tested and deployed the Teleo-Reactive EXecutive
(\texttt{T-REX}) an on-board adaptive control system that integrates
AI based planning and probabilistic state estimation in a hybrid
executive \cite{mcgann08a,mcgann08b,Py10}.  Probabilistic state
estimation integrates a number of science observations to produce a
likelihood that the vehicle sensors perceive a feature of
interest. Onboard planning and execution enables adaptation of
navigation and instrument control based on the probability of having
detected such a phenomenon. It further enables goal-directed
commanding within the context of projected mission state and allows
for replanning for off-nominal situations and opportunistic science
events. The framework in addition to being used on an AUV, is general
enough be used for controlling a personal robot \cite{mcgann2009} and
is deployed on a European planetary rover testbed
\cite{goac11}. Probabilistic state estimation is discussed elsewhere
\cite{mcgann08d} and out of scope for this chapter. Deliberation and
reaction are integrated systematically over different temporal and
functional scopes within a single agent and a single model that covers
the needs of high-level mission management, low-level navigation,
instrument control, and detection of unstructured and poorly
understood phenomena. \rx is deployed on MBARI's Dorado AUV shown in
Fig. \ref{fig:auv-fig}, which to the best of our knowledge is the only
operational marine platform anywhere being used for routine scientific
surveys with onboard plan synthesis.

\begin{figure}[t]
  \centering
  \vskip-5pt
  \includegraphics[scale=0.1]{figs/MBARI-AUV.jpg}
  \caption{\small The MBARI \emph{Dorado} AUV on its support vessel
    the R/V \emph{Zephyr}.}
  \label{fig:auv-fig}
  \vskip-0.3cm
\end{figure}

This chapter is organized as follows. We start with some foundational
concepts in our Planning framework in Section
\ref{sec:concepts}. \comment{add others} We conclude with Section
\ref{sec:conclusion}.
